---
title: Latitude and Longitude Coordinates
---
Use the [`mapview` package](https://r-spatial.github.io/mapview/index.html) to make interactive maps.  In this example, georeference Starbucks coffee shop locations in North Carolina (2012). ^[[5 Visualizations in 5 Minutes](http://www.computerworld.com/article/2893271/business-intelligence/5-data-visualizations-in-5-minutes-each-in-5-lines-or-less-of-r.html). ComputerWorld.com by Sharon Machlis]


## Load Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(mapview)
```

## Load Data

2012 Starbucks locations ([data source](https://github.com/libjohn/mapping-with-R/blob/master/data/All_Starbucks_Locations_in_the_US_-_Map.csv))

```{r load-data, message=FALSE, warning=FALSE}
starbucks <- read_csv("https://raw.githubusercontent.com/libjohn/mapping-with-R/master/data/All_Starbucks_Locations_in_the_US_-_Map.csv")
```

### Subset Data to North Carolina
```{r filter-dataset, message=FALSE, warning=FALSE}
starbucksNC <- starbucks  %>% 
  filter(State == "NC")

starbucksNC %>% 
  glimpse()
```

## Make the Map

In this example, plot latitude (y coordinates) and longitude (x coordinates).  Then set the map projection^[[Projection / CRS](https://guides.library.duke.edu/r-geospatial/CRS)] to a common projection standard such as _WGS84_ via the argument `crs = 4326`.)

```{r map1}
mapview(starbucksNC, xcol = "Longitude", ycol = "Latitude", crs = 4269, grid = FALSE)
```

## Alternative: Transform data to Spatial object

Another way to plot the x & y coordinates is by transforming the starbucksNC tibble (i.e. the starbucksNC data frame) into a spacial data frame via the simple features function, `st_as_sf()`. Again, set the [map projection](https://en.wikipedia.org/wiki/Map_projection) to a common standard such as _WGS84_ via the `crs=` argument.

```{r convert2spacial}
sbux_sf <- st_as_sf(starbucksNC, coords = c("Longitude", "Latitude"),  crs = 4326)
```

### Now Map the sf object.

Below, you can plot the latitude and longitude coordinate, and set the `map.types` argument to `openStreetMap`.  Base maps are set with `map.types`.  For example a high contrast, black and white basemap can be set with the argument `map.types = "Stamen.Toner"`. See available [map types](http://leaflet-extras.github.io/leaflet-providers/preview/), or leave out the map.types argument for a set of default base maps to choose via the layering button.

```{r plotmap}
#mapview(sbuxsfc, map.types = "Stamen.Toner")
mapview(sbux_sf)
```







