---
title: ggmap -- Latitude and Longitude Coordinates
output: html_document
#  html_document:
#    toc: FALSE
---
&nbsp;

In this exmaple, build a static map quickly and easily using `ggmap`. The `ggmap` package enables the integration of popular raster base-maps with ggplot2 syntax.  This example


## Load Packages

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggmap)
```

## Load Data

2012 Starbucks locations

```{r load-data, message=FALSE, warning=FALSE}
starbucks <- read_csv("data/All_Starbucks_Locations_in_the_US_-_Map.csv")
```

### Subset Data to North Carolina

Using the same Starbucks from the previous example...
```{r filter-dataset}
starbucksNC <- starbucks  %>% 
  filter(State == "NC")
```

## ggmap

### Indentify base map and mapping location

Some of the most popular base maps come from the Google Maps source.  Using ggmap, the following maptypes are available from the Google Maps source ...

- terrain
- satellite
- roadmap
- hybrid

```{r message=FALSE, warning=FALSE}
get_local_spot <-  get_map("Durham North Carolina", maptype = "roadmap", zoom = 10) 
```

Alternatively, use `help(get_map)` to identify additional map sources and other arguments for `maptype`.  For example...

- Stamen maps: "terrain", "watercolor", and "toner" 
- or a positive integer for cloudmade maps (see ?get_cloudmademap)

### Display Base Map

```{r}
ggmap(get_local_spot)
```

### Display points with geom_point

```{r message=FALSE, warning=FALSE}
ggmap(get_local_spot) + 
  geom_point(data = starbucksNC, aes(x = Longitude, y = Latitude), color = "navy", size = 1)
```

### Frame the Map

Altering the map frame from a square to a rectangle requires knowing the coordinates of the Southwest and Northeast corners of the map frame.  This time, display a different basemapusing the the Stamen source and the Toner maptype.

```{r message=FALSE, warning=FALSE}
mapbox <- c(-80, 35.6, -78.5, 36.25)
piedmont_coffee <- get_map(location = mapbox, source = "stamen", maptype = "toner", zoom = 9)
```


```{r}
ggmap(piedmont_coffee)
```

#### Plot Coffee Shops

ggplot2 syntax comes in handy for creating various map views.


```{r message=FALSE, warning=FALSE}

ggmap(piedmont_coffee) +
  geom_point(data = starbucksNC, 
             aes(x = Longitude, y = Latitude), 
             color = "red", 
             size = 2,
             alpha = 0.7) +
  theme(axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      rect = element_blank(),
      axis.title.y=element_blank(),
      axis.title.x=element_blank())
```


