---
title: Latitude and Longitude Coordinates
output: html_document
#  html_document:
#    toc: FALSE
---
&nbsp;

To introduce the [`mapview` package](https://r-spatial.github.io/mapview/index.html), make an interactive map with Starbucks coffee shop locations in North Carolina (2012).  (Adapted from Machlis.)^[[5 Visualizations in 5 Minutes](http://www.computerworld.com/article/2893271/business-intelligence/5-data-visualizations-in-5-minutes-each-in-5-lines-or-less-of-r.html). ComputerWorld.com by Sharon Machlis]


## Load Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(leaflet)
library(mapview)
```

## Load Data

2012 Starbucks locations ([data source](https://github.com/libjohn/mapping-with-R/blob/master/data/All_Starbucks_Locations_in_the_US_-_Map.csv))

```{r load-data, message=FALSE, warning=FALSE}
starbucks <- read_csv("data/All_Starbucks_Locations_in_the_US_-_Map.csv")
```

### Subset Data to North Carolina
```{r filter-dataset, message=FALSE, warning=FALSE}
starbucksNC <- starbucks  %>% 
  filter(State == "NC")

starbucksNC
```

## Transform to Spatial Data

Transform our standard tibble (i.e. the starbucksNC data frame) into a spacial data frame via the simple features function, `st_as_sf()`. Here we set the [map projection](https://en.wikipedia.org/wiki/Map_projection) to a common standard, WGS84, via the argument `crs = 4326`.)

```{r convert2spacial}
sbuxsfc <- st_as_sf(starbucksNC, coords = c("Longitude", "Latitude"),  crs=4326)
```

## Visualize the Map

Below, you can plot the latitude and longitude coordinate, and set the `map.types` argument to `openStreetMap`.  Base maps are set with `map.types`.  For example a high contrast, black and white basemap can be set with the argument `map.types = "Stamen.Toner"`. See available [map types](http://leaflet-extras.github.io/leaflet-providers/preview/), or leave out the map.types argument for a set of default base maps to choose via the layering button.

```{r plotmap}
#mapview(sbuxsfc, map.types = "Stamen.Toner")
mapview(sbuxsfc)
```



## Save the Map

You can save a map as an interactive HTML page or a static `.png, .pdf, or .jpeg` file with the [`mapshot` function](https://r-spatial.github.io/mapview/reference/mapshot.html).  For example:

- `mapshot(starNCmap, file = "map.png")`
- `mapshot(starNCmap, url = "map.html")`


