---
title: "TidyCensus"
output: html_document
---


Goal:  Make a choropleth with census data

- define choropleth
- define polygons
- define census geography
- Briefly Introduce ACS and decennial census 

    - census key 


```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidycensus)
library(leaflet)
library(sf)
library(stringr)
```


## Create a Simple Features dataframe using TidyCensus
```{r load-data, message=FALSE, warning=FALSE}
nc_pop <- 
  get_acs(geography = "county",
          variables = "B01003_001",
          state = "NC",
          geometry = TRUE)

nc_pop
```




## Make Choropleth in Leaflet

Generate a color palette for the colors and legend

```{r colorpalette}
MapPalette <- colorQuantile(palette = "viridis", domain = nc_pop$estimate, n = 10)
```


Make choropleth by filling county polygons (census geography) with correlated value (population variable *B01003_001*)

`st_transform()` - converts coordinate system of simple features

```{r make_choropleth}
nc_pop %>%
    st_transform(crs = "+init=epsg:4326") %>%
    leaflet(width = "100%") %>%
    addProviderTiles(provider = "CartoDB.Positron") %>%
    addPolygons(popup = ~ str_extract(NAME, "^([^,]*)"),
                stroke = FALSE,
                smoothFactor = 0,
                fillOpacity = 0.7,
                color = ~ MapPalette(estimate)) %>%
    addLegend("bottomright", 
              pal = MapPalette, 
              values = ~ estimate,
              title = "Population percentiles",
              opacity = 1)
```

## Add another layer

Add locations of the the 2012 Starbucks stores.  The Starbucks locations were generated and plotted in the previous exercise.  Re-running exercise 1 will ensure the starbucks$NC object is in your R environment.

```{r add_layer}
nc_pop %>%
    st_transform(crs = "+init=epsg:4326") %>%
    leaflet(width = "100%") %>%
    addProviderTiles(provider = "CartoDB.Positron") %>%
    addPolygons(popup = ~ str_extract(NAME, "^([^,]*)"),
                stroke = FALSE,
                smoothFactor = 0,
                fillOpacity = 0.7,
                color = ~ MapPalette(estimate)) %>%
    addLegend("bottomright", 
              pal = MapPalette, 
              values = ~ estimate,
              title = "Population percentiles",
              opacity = 1) %>% 
    addTiles() %>% 
  # The next paragraph adds a layer from exercise 01
    addCircleMarkers(data = starbucksNC, 
               lat = ~ Latitude, 
               lng = ~ Longitude, 
               popup = starbucksNC$Name,
               weight = 1,
               radius = 0.6,
               opacity = 0.5, 
               color = "black")
```

##  Census

During the workshop, and in the video recording, I will discuss the following concepts in more detail.

- ACS v Decennial
- Variable Names / Numbers
- More on Census Geography (shapefiles)

### Variables

The Census is a very large collection of data.  Many casual users of Census data are interested in a single data point, for example population by County.  Given the complexity and richness of available Census data, finding a useful data variable can be quite a bit of work.  The links below are a some methods one might use to associate and identify the proper code name with a Census data variable.

- [Searching for Variables](https://walkerke.github.io/tidycensus/articles/basic-usage.html#searching-for-variables)
- See Also [Tiger Metadata](https://www2.census.gov/geo/tiger/TIGER_DP/2015ACS/Metadata/STATE_METADATA_2015.txt)
- See Also [SocialExplorer Tables](https://www.socialexplorer.com/data/ACS2015_5yr/metadata/?ds=ACS15_5yr)

## Shapefiles

In the [*Basic Thematic Mapping* section](031_thematic_mapping.html) of the [Simple Features module](030_thematic_sf.html) we will use the `tigris` package to gather State level, Census geography shapefiles.  Shapefiles are an important GIS data standard.  If you need shapefiles for other geographies, please consult our [*GIS Data: Starting Points* guide](https://guides.library.duke.edu/gisdata) --  a very handy collection of GIS data sources, much of which will be in the shapefiles format.  

## End Notes

This session based on 

- Kyle Walker's [TidyCensus](https://walkerke.github.io/tidycensus/) package
    
- Julia Silge's [leaflet implementation of tidycensus](https://juliasilge.com/blog/using-tidycensus/)