---
title: "tidycensus"
output: html_document
---

The [tidycensus](https://walkerke.github.io/tidycensus/index.html) package, developed by [Kyle Walker](https://walkerke.github.io/), is very **convenient and easy to use package for making choropleth maps** from United States Department of **Census data**, specifically from the Decennial and ACS Census reports.  This package makes it possible to gather census variables and conveniently join those variables with "Census Geography" (i.e. aka "shapefiles", or polygons.)  Visualization, or plotting, maps can be done with separate packages.  


```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(stringr)
library(tidycensus)
library(sf)
library(leaflet)
library(mapview)
```

## Census API Key

Use the tidycensus package to gather Census data and join that data with Census geography (i.e. geometry, i.e. shapefiles, i.e. polygons).  First, you will need to get a [free Census API key](https://api.census.gov/data/key_signup.html).  Kyle Walker's [*Basic usage of tidycensus*](https://walkerke.github.io/tidycensus/articles/basic-usage.html) documents this process. 

```{}
census_api_key("YOUR API KEY GOES HERE")
```

### .Renviron File

See also Kyle's [more detailed documentation](https://walkerke.github.io/tidycensus/reference/census_api_key.html) for putting the key into your environment file.  But skip that for the moment.


## TidyCensus -- Get Data

Create a Simple Features dataframe using `tidycensus::get_acs()`

The Census population variable we'll use is "B01003_001".  More information about identifying Census variables is available at the [bottom of this page](#variables).
```{r load-data, message=FALSE, warning=FALSE, include=FALSE}
nc_pop <- 
  get_acs(geography = "county",
          variables = "B01003_001",
          state = "NC",
          geometry = TRUE)

#nc_pop
```
```{}
nc_pop <- 
  get_acs(geography = "county",
          variables = "B01003_001",
          state = "NC",
          geometry = TRUE)
```




## Make Choropleth via mapview

Identify which variable will be used to create the color ramp shading.  Assign this variable with the `zcol` argument.  The `estimate` variable was extracted via the `tidycensus::get_acs()` function.


```{r make_choropleth}
mapview(nc_pop, zcol = "estimate")
```


## Add another layer

Now we'll geolocate the Starbucks stores and add those locations as a layer over our choropleth.  The Starbucks locations were generated and plotted in the previous exercise.  Here we regenerate the StarbuckNC object.

### Load Lat/Long Data

```{r load-data_02, message=FALSE, warning=FALSE}
starbucks <- read_csv("data/All_Starbucks_Locations_in_the_US_-_Map.csv")
```


Subset Starbucks Data to North Carolina
```{r filter-dataset}
starbucksNC <- starbucks  %>% 
  filter(State == "NC")
```

Convert the `starbuckNC` dataframe to a spatial (sf) object and assign the same projection as the `nc_pop` spatial object.

```{r}
starbucksNC <- st_as_sf(starbucksNC, coords = c("Longitude", "Latitude"),  crs = st_crs(nc_pop))
```



Generate the map with multiple layers.  You can read more about additional arguments such as `homebutton, legend, alpha, cex` in the [`mapview()` documentation](https://r-spatial.github.io/mapview/reference/mapView.html).  Read about the many more mapview functions in the [full documentation](https://r-spatial.github.io/mapview/reference/index.html).

```{r}
mymap <- mapview(nc_pop, 
                 zcol = "estimate", 
                 homebutton = FALSE) + 
  mapview(starbucksNC, 
          zcol = "Name", 
          legend = FALSE, 
          alpha = 0.5, cex = 3, 
          homebutton = FALSE) 

addLogo(mymap, "images/Rfun3.png",
        position = "bottomright",
        offset.x = 8,
        offset.y = 38,
        width = 100,
        height = 100)
```

##  Census

During the workshop, and in the video recording, I will discuss the following concepts in more detail.

- ACS v Decennial
- Variable Names / Numbers
- More on Census Geography (shapefiles)

### Variables

The Census is a very large collection of data.  Many casual users of Census data are interested in a single data point, for example population by County.  Given the complexity and richness of available Census data, finding a useful data variable can be quite a bit of work.  The links below are a some methods one might use to associate and identify the proper code name with a Census data variable.

- [Searching for Variables](https://walkerke.github.io/tidycensus/articles/basic-usage.html#searching-for-variables)
- See Also [Tiger Metadata](https://www2.census.gov/geo/tiger/TIGER_DP/2015ACS/Metadata/STATE_METADATA_2015.txt)
- See Also [SocialExplorer Tables](https://www.socialexplorer.com/data/ACS2015_5yr/metadata/?ds=ACS15_5yr)

## Shapefiles

In the [*tmap* section](031_thematic_mapping.html) of the [Simple Features module](030_thematic_sf.html) we will use the `tigris` package to gather State level, Census geography shapefiles.  Shapefiles are an important GIS data standard.  There are many other standards, although shapefiles have a very broad user base.  If you need shapefiles for other geographies, please consult our [*GIS Data: Starting Points* guide](https://guides.library.duke.edu/gisdata) --  a very handy collection of GIS data sources, much of which will be in the shapefiles format.  

## End Notes

This session based on 

- Kyle Walker's [TidyCensus](https://walkerke.github.io/tidycensus/) package