---
title: "TidyCensus"
output: html_notebook
---


Goal:  Make a choropleth with census data

- define choropleth
- define poloygons
- Briefly Introduce ACS census and decennial census 

    - census key 



```{r}
library(tidyverse)
library(tidycensus)
library(leaflet)
library(sf)
library(stringr)
```


This session based on 

- Kyle TidyCensus

    - [Searching for Variables](https://walkerke.github.io/tidycensus/articles/basic-usage.html#searching-for-variables)
    - See Also [Tiger Metadata](https://www2.census.gov/geo/tiger/TIGER_DP/2015ACS/Metadata/STATE_METADATA_2015.txt)
    - See Also [SocialExplorer Tables](https://www.socialexplorer.com/data/ACS2015_5yr/metadata/?ds=ACS15_5yr)
& 
- Silge's [leaflet implementation of tidycensus](https://juliasilge.com/blog/using-tidycensus/)

### Create a Simple Features dataframe using TidyCensus
```{r}
nc_pop <- 
  get_acs(geography = "county",
          variables = "B01003_001",
          state = "NC",
          geometry = TRUE)

nc_pop
```




## Make Coropleth in Leaflet

Generate a pallet for the colors and legend

```{r}
MapPalette <- colorQuantile(palette = "viridis", domain = nc_pop$estimate, n = 10)
```


Make Chropleth by filling county polygons (census geography) with correlated value

`st_transform()` - converts coordinate system of simple features

```{r}
nc_pop %>%
    st_transform(crs = "+init=epsg:4326") %>%
    leaflet(width = "100%") %>%
    addProviderTiles(provider = "CartoDB.Positron") %>%
    addPolygons(popup = ~ str_extract(NAME, "^([^,]*)"),
                stroke = FALSE,
                smoothFactor = 0,
                fillOpacity = 0.7,
                color = ~ MapPalette(estimate)) %>%
    addLegend("bottomright", 
              pal = MapPalette, 
              values = ~ estimate,
              title = "Population percentiles",
              opacity = 1)
```



```{r}
nc_pop %>%
    st_transform(crs = "+init=epsg:4326") %>%
    leaflet(width = "100%") %>%
    addProviderTiles(provider = "CartoDB.Positron") %>%
    addPolygons(popup = ~ str_extract(NAME, "^([^,]*)"),
                stroke = FALSE,
                smoothFactor = 0,
                fillOpacity = 0.7,
                color = ~ MapPalette(estimate)) %>%
    addLegend("bottomright", 
              pal = MapPalette, 
              values = ~ estimate,
              title = "Population percentiles",
              opacity = 1) %>% 
    addTiles() %>% 
    addCircleMarkers(data = starbucksNC, 
               lat = ~ Latitude, 
               lng = ~ Longitude, 
               popup = starbucksNC$Name,
               weight = 1,
               radius = 0.6,
               opacity = 0.5, 
               color = "black")
```

##  Census

- ACS v Decennial
- Variable Names / Numbers
- More on Census Geography

