---
title: "Thematic Mapping"
output: html_notebook
---

from:  https://www.computerworld.com/article/3175623/data-analytics/mapping-in-r-just-got-a-whole-lot-easier.html


```{r}
library(tidyverse)       # Tidyverse for Tidy Data
library(readxl)
library(tmap)            # Thematic Mapping
library(tmaptools) 
library(tigris)          # Get Census Geography Poloygons
library(sf)
```


## Get shapefiles in a "simple features" format

```{r}
us_geo <- states(class = "sf")
# us_geo_spdf <- states()  # spdf uses @data slots -- old school
```

## Structure of a sf file

```{r}
class(us_geo)
glimpse(us_geo)
```

## Look at the sf in a data frame
```{r}
#as_tibble(us_geo)
```

## Get BLS Data

https://data.bls.gov/oes/#/occGeo/One%20occupation%20for%20multiple%20geographical%20areas

-  One occupation for multiple geographical areas

    - Mental Health and Substance Abuse Social Workers
    
        - State
        
        - All States in this list
        
        - Annual Mean wage
        
            - Excel
            
- Read the Data in with the RStudio "Import Dataset" wizard available in the *Environment* tab.  This will generate the code below and ensure the import

    - Skips the first 4 lines
    - Coerces the 2nd column to numeric

```{r}
Salary4Helpers <- 
  read_excel("data/OES_Report.xlsx",
             col_types = c("text", "numeric"), 
             skip = 4)

#Salary4Helpers
```

        
        

## Wrangle the Data

```{r}
BlsWage_ToJoin <- Salary4Helpers %>% 
  rename(Area = "Area Name") %>% 
  rename(wages = "Annual mean wage(2)") %>% 
  mutate(State = gsub("\\(\\d{7}\\)", "", Area)) %>% 
  filter(wages != "NA_character_") %>% 
  select(State, wages)

#BlsWage_ToJoin
```


## Append Data

Using the `append_data()` function of the `tmaptools` package, append BLS data to the previously loaded shape object

```{r}
HelperShapeObject <- append_data(us_geo, BlsWage_ToJoin, 
                                 key.shp = "NAME",
                                 key.data = "State")
#as_tibble(HelperShapeObject)
```


## Quick Thematic Map

```{r}
qtm(HelperShapeObject, fill = "wages")
```


## Contiguous 48 states

Filter to only the contiguous 48 states + D.C.

```{r}
contiguous_states <- HelperShapeObject %>% 
  filter(REGION != 9) %>% 
  filter(STUSPS != "AK") %>% 
  filter(STUSPS != "HI")
  
```


## Make Choropleth

```{r}
tm_shape(contiguous_states) +
  tm_polygons("wages", id = "Name")
```


## Projection

Mark likes the USA_Contiguous_Albers_Equal_Area_Conic_USGS_version projection for the continental US. `EPSG:5070`

```{r}
contiguous_states %>% 
  st_transform(5070) %>% 
  tm_shape() +
  tm_polygons("wages", id = "Name")
```


### Alternative Syntax


```{r}
tm_shape(contiguous_states, projection = 5070) +
  tm_polygons("wages", id = "Name")
```

### Explore `tmap` syntax and functions 


```{r}
tm_shape(contiguous_states, projection = 5070) +
  tm_borders(col = "black", alpha = 0.4) +
  tm_fill(col = "REGION", alpha = 0.6) +
  tm_style_beaver(title = "Regions of the USA", 
                  attr.color = "navy", 
                  title.position = c("center", "top"), 
                  title.bg.color = "yellow")
```


