---
title: "Thematic Mapping with tmap"
output: html_document
---


```{r load-libraries, message=FALSE, warning=FALSE}
library(tidyverse)       # Tidyverse for Tidy Data
library(readxl)
library(tmap)            # Thematic Mapping
library(tmaptools) 
library(tigris)          # Get Census Geography Poloygons
library(sf)
```


## Shapefiles as sf 

Using the `tigris` package, get Census Tiger shapefiles for census geographies.  Tigris will return the shapefile in the `sf`, or simple features, format.  (See also, [more information on Shapefiles](02_choropleth.html#shapefiles).)

```{}
us_geo <- tigris::states(class = "sf")
# us_geo_spdf <- states()  # spdf uses @data slots -- old school
```

```{r get-data, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
us_geo <- tigris::states(class = "sf")
# us_geo_spdf <- states()  # spdf uses @data slots -- old school
```

## Data Structure 

The advantage of managing your spatial data as "Simple Features" (i.e. sf) over `sp` is that `sf` data can easily be viewed and manipulated as a rectangular data frame, before visualizing.   `sp` is the predecessor to `sf`, but uses the old school @data slots method.  If you're used to this it can be a highly functional data structure.  But if you're inclined toward the tidyverse method, you'll see your data in a more familiar and easy-to-view format.  Below are two methods of viewing the structure of the downloaded shapefiles.

```{r data-structure}
class(us_geo)
glimpse(us_geo)
```

## sf as data frame

And as noted, here's the data frame view.  Notice the geometry (polygon shape) in the far-right column of the data frame.


```{r see-tibble}
as_tibble(us_geo)
```

## Get BLS data

I've already downloaded and stored some data from the Bureau of Labor Statistics.  Those data are stored in an excel file in the `data` directory of the [repository](https://github.com/libjohn/mapping-with-R):  `data/OES_Report.xlsx`.  **The goal is to attach this data to the previously downloaded shapefiles.**  

But you may be interested in how I gathered the data. below are some summary notes documenting my steps of gathering the data from the Bureau of Labor Statistics.  

https://data.bls.gov/oes/#/occGeo/One%20occupation%20for%20multiple%20geographical%20areas

-  One occupation for multiple geographical areas

    - Mental Health and Substance Abuse Social Workers
    
        - State
        
        - All States in this list
        
        - Annual Mean wage
        
            - Excel
            
- Read the Data in with the RStudio "Import Dataset" wizard available in the *Environment* tab.  This will generate the code below and ensure the import

    - Skips the first 4 lines
    - Coerces the 2nd column to numeric

```{r get-blsdata, message=FALSE, warning=FALSE}
Salary4Helpers <- 
  read_excel("data/OES_Report.xlsx",
             col_types = c("text", "numeric"), 
             skip = 4)

Salary4Helpers
```

        
        

## Wrangle the data

Before we join the BLS data to the shapefile we need to transform the structure of the downloaded BLS data

```{r wrangle_bls_data, message=FALSE, warning=FALSE}
BlsWage_ToJoin <- Salary4Helpers %>% 
  rename(Area = "Area Name") %>% 
  rename(wages = "Annual mean wage(2)") %>% 
  mutate(State = gsub("\\(\\d{7}\\)", "", Area)) %>% 
  filter(wages != "NA_character_") %>% 
  select(State, wages)

#BlsWage_ToJoin
```


## Append data

Using the `append_data()` function of the `tmaptools` package, append BLS data to the previously loaded shape object

```{r make_nice_df, message=FALSE, warning=FALSE}
HelperShapeObject <- append_data(us_geo, BlsWage_ToJoin, 
                                 key.shp = "NAME",
                                 key.data = "State")
as_tibble(HelperShapeObject)
```


## Quick Thematic Map

```{r make_qtm}
qtm(HelperShapeObject, fill = "wages")
```


## Contiguous 48 states

Filter to only the contiguous 48 states + D.C.

```{r subset-data}
contiguous_states <- HelperShapeObject %>% 
  filter(REGION != 9) %>% 
  filter(STUSPS != "AK") %>% 
  filter(STUSPS != "HI")
  
```


## Make Choropleth

```{r make-choropleth}
tm_shape(contiguous_states) +
  tm_polygons("wages", id = "Name")
```


## Projection

Mark likes the USA_Contiguous_Albers_Equal_Area_Conic_USGS_version projection for the continental US. `EPSG:5070`

```{r fix-projection}
contiguous_states %>% 
  st_transform(5070) %>% 
  tm_shape() +
  tm_polygons("wages", id = "Name")
```


### Alternative Syntax


```{r alternative-syntax}
tm_shape(contiguous_states, projection = 5070) +
  tm_polygons("wages", id = "Name")
```

### Explore `tmap` syntax and functions 


```{r tmap-syntax}
tm_shape(contiguous_states, projection = 5070) +
  tm_borders(col = "black", alpha = 0.4) +
  tm_fill(col = "REGION", alpha = 0.6) +
  tm_style_beaver(title = "Regions of the USA", 
                  attr.color = "navy", 
                  title.position = c("center", "top"), 
                  title.bg.color = "yellow")
```



## End Notes

This session inspired by  https://www.computerworld.com/article/3175623/data-analytics/mapping-in-r-just-got-a-whole-lot-easier.html
