---
title: "Thematic Mapping with geom_sf"
output: html_document
---


```{r load-libraries, message=FALSE, warning=FALSE}
library(tidyverse)       # Tidyverse for Tidy Data
library(readxl)
library(tigris)
library(sf)
library(viridis)
library(ggplot2) 
```


## Shapefiles as sf 

Repeating steps from the [previous section](031_thematic_mapping.html), using the `tigris` package, get Census Tiger shapefiles for census geographies.  

```{}
us_geo <- tigris::states(class = "sf")
```

## Get BLS data

As mentioned before, the data are from the Bureau of Labor Statistics.  These data are stored in an excel file in the `data` directory of the [repository](https://github.com/libjohn/mapping-with-R):  `data/OES_Report.xlsx`.  

Again, from the previous section you will see how these data were gathered, loaded, transformed, and joined. 

```{}
Salary4Helpers <- 
  read_excel("data/OES_Report.xlsx",
             col_types = c("text", "numeric"), 
             skip = 4)

Salary4Helpers
```

## Wrangle the data

As before...  

```{}
BlsWage_ToJoin <- Salary4Helpers %>% 
  rename(Area = "Area Name") %>% 
  rename(wages = "Annual mean wage(2)") %>% 
  mutate(State = gsub("\\(\\d{7}\\)", "", Area)) %>% 
  filter(wages != "NA_character_") %>% 
  select(State, wages)
```


## Append data

As before, using the `append_data()` function of the `tmaptools` package, append BLS data to the previously loaded shape object

```{}
HelperShapeObject <- append_data(us_geo, BlsWage_ToJoin, 
                                 key.shp = "NAME",
                                 key.data = "State")
as_tibble(HelperShapeObject)
```


## Contiguous 48 states

As before, filter to only the contiguous 48 states + D.C.

```{}
contiguous_states <- HelperShapeObject %>% 
  filter(REGION != 9) %>% 
  filter(STUSPS != "AK") %>% 
  filter(STUSPS != "HI")
  
```



## Projection

Using the USA_Contiguous_Albers_Equal_Area_Conic_USGS_version projection for the continental US. `EPSG:5070`

```{}
contiguous_states %>% 
  st_transform(5070) %>% 
  tm_shape() +
  tm_polygons("wages", id = "Name")
```


## ggplot2 with geom_sf and viridis

In this section we introduce making shapefiles with ggplot2.  ggplot2 is one of the more popular and broadly distributed graphics packages used in the R community. I also reveresed the direction of the color scale.  After consulting with my visualization colleagues it seems this may have been a non-standard action on my part.  But I leave the `direction` argument here for the curious.

```{r with-ggplot-geom_sf}
contiguous_states %>% 
  ggplot(aes(fill = wages, color = wages)) +
  geom_sf() +
  coord_sf(crs = 5070) +
  scale_fill_viridis(option = "viridis", direction = -1) + 
  scale_color_viridis(option = "viridis", direction = -1)


```


## End Notes

This session inspired by  https://www.computerworld.com/article/3175623/data-analytics/mapping-in-r-just-got-a-whole-lot-easier.html
