---
title: "Faceted Thematic Mapping"
---

```{r load-libraries, message=FALSE, warning=FALSE}
library(tidyverse)       # Tidyverse for Tidy Data
library(readxl)
library(tigris)          # Get Census Geography Poloygons
library(sf)
library(tidycensus)
```

## Shapefiles as sf 

Using the `tigris` package get Census Tiger shapefiles for census geographies.  Tigris will return the shapefile in the `sf`, or simple features, format.

```{r get-data, message=FALSE, warning=FALSE}
us_geo <- tigris::states(class = "sf") %>% 
  shift_geometry()
```

## Get BLS data

I've already downloaded and stored some data from the Bureau of Labor Statistics.  Thoses data are stored in an excel file in the `data` directory of the repository:  `data/OES_Report.xlsx`.  **The goal is to attach this data to the previously downloaded shapefiles.**  

But you may be intersted in how I gathered the data. below are some summary notes documenting my steps of gathering the data from the Bureau of Labor Statistics.  

https://data.bls.gov/oes/#/occGeo/One%20occupation%20for%20multiple%20geographical%20areas

-  One occupation for multiple geographical areas

    - Mental Health and Substance Abuse Social Workers  (Also, Secondary School Teacher, Waiter, Legislator, and Paralegals)
    
        - State
        
        - All States in this list
        
        - Annual Mean wage
        
            - Excel
            
### Import the Data into R 
    
```{r message=FALSE, warning=FALSE}
my_xl_files <- fs::dir_ls(path = "data", glob = "*.xlsx")

my_df <- my_xl_files %>% 
  map_dfr(read_excel, 
          col_types = c("text", "numeric"), 
          skip = 4, 
          .id = "sheet")
```

```{r}
state_names <- us_geo %>% 
  filter(REGION != "9") %>% 
  select(NAME) %>% 
  st_drop_geometry() 
```


### Wrangle the data

Before we join the BLS data to the shapefile `us_geo` we need to transform BLS data

```{r}
my_df <- my_df %>% 
  rename(area = "Area Name",
         wages = "Annual mean wage(2)",
         type = sheet) %>% 
  mutate(State = str_extract(area, '.*(?=\\()')) %>% 
  mutate(type = str_extract(type, "(?<=data/OES_)\\w+")) %>% 
  filter(State != "Guam")
```

### Missing data

Some of the BLS data are missing making it hard to visualize.  As a remedy to this problem we will wrangle the data by preserving shape geometry even for states without any wage data.

```{r}
missing_states_legislators <- state_names %>% 
  anti_join(my_df %>% filter(type == "legislator"), 
            by = c("NAME" = "State")) %>% 
  mutate(type = "legislator") %>% 
  rename(State = NAME)

missing_states_legislators
```

```{r}
my_df <- my_df %>% 
  bind_rows(missing_states_legislators)
```

### Join data

Using the `dplyr::left_join` merge BLS data to the previously loaded shape object

```{r}
my_df <- us_geo %>% 
  left_join(my_df, by = c("NAME" = "State"))
```

### 50 states

Filter to the 50 states + D.C.

```{r}
my_df <- my_df %>% 
  filter(REGION != 9) 
```

## Get Population Data -- Tidycensus

### identify and pick census variables

B01003_001E = Total Population
B06011_001E = Median income in the past 12 months

```{r identify_ACS_variables}
variables_census <- load_variables(2015, "acs5", cache = TRUE)
```

### Get Median Income

```{r get_median_income}
us_pop <- 
  get_acs(geography = "state",
          variables = "B06011_001E",
          geometry = TRUE)  %>% 
  shift_geometry()

us_pop <- us_pop %>% 
  mutate(type = "USA Median Income") %>% 
  rename(wages = estimate) %>% 
  filter(NAME != "Puerto Rico")
```

### Append Census Data

```{r}
my_df <- bind_rows(my_df, us_pop)
```

### More Munging

Make the category variable a categorical factor with levels.  This will improve the order of the facets when displayed.

```{r}
display_levels <- c("USA Median Income", "Legislator", 
                    "Paralegals", "Substance Abuse Counselor", 
                    "Teacher", "Waiters")

my_df <- my_df %>% 
  mutate(category = case_when(
    type == "USA Median Income" ~ "USA Median Income",
    type == "legislator" ~ "Legislator",
    type == "paralegals" ~ "Paralegals",
    type == "Report" ~ "Substance Abuse Counselor",
    type == "secondary_school_teacher" ~ "Teacher",
    type == "waiters" ~ "Waiters"
  )) %>% 
  mutate(category = factor(category, display_levels))
```

### Display Map

```{r}
my_df %>%
  ggplot(aes(fill = wages, color = wages)) +
  geom_sf() +
  coord_sf(crs = 5070, datum = NA) +
  scale_fill_viridis_c() + 
  scale_color_viridis_c() +
  facet_wrap(~ category, 
             nrow = 3, ncol = 2)
```



### Save Map

```{r save_it}
ggsave("facet_map.png", width = 8, height = 8, units = "in")
```
