---
title: "Thematic Mapping"
output: html_notebook
---

from:  https://www.computerworld.com/article/3175623/data-analytics/mapping-in-r-just-got-a-whole-lot-easier.html


```{r}
library(tidyverse)       # Tidyverse for Tidy Data
library(readxl)
library(tmap)            # Thematic Mapping
library(tmaptools) 
library(tigris)          # Get Census Geography Poloygons
library(sf)
```


# Get Shapfiles in a "simple features format"

```{r}
us_geo <- states(class = "sf")
# us_geo_spdf <- states()  # spdf uses @data slots -- old school
```

# Structure of a sf file

```{r}
class(us_geo)
glimpse(us_geo)
```

## Look at the Tibble
```{r}
#as_tibble(us_geo)
```

## Get BLS Data

https://data.bls.gov/oes/#/occGeo/One%20occupation%20for%20multiple%20geographical%20areas

-  One occupation for multiple geographical areas

    - Mental Health and Substance Abuse Social Workers
    
        - State
        
        - All States in this list
        
        - Annual Mean wage
        
            - Excel
            
- Read the Data in with the readr data wizard

```{r}
# Skip first four lines
# Make second column numeric

Salary4Helpers <- 
  read_excel("~/R/github/mapping-with-R/data/OES_Report.xlsx", 
    col_types = c("text", "numeric"), skip = 4)

#Salary4Helpers
```

        
        

# Wrangle the Data

```{r}
BlsWage_ToJoin <- Salary4Helpers %>% 
  rename(Area = "Area Name") %>% 
  rename(wages = "Annual mean wage(2)") %>% 
  mutate(State = gsub("\\(\\d{7}\\)", "", Area)) %>% 
  filter(wages != "NA_character_") %>% 
  select(State, wages)

#BlsWage_ToJoin
```


## Append Data

Using the `append_data()` function of the `tmaptools` package, append BLS data to the previously loaded shape object

```{r}
HelperShapeObject <- append_data(us_geo, BlsWage_ToJoin, 
                                 key.shp = "NAME",
                                 key.data = "State")
#as_tibble(HelperShapeObject)
```


## Quick Thematic Map

```{r}
qtm(HelperShapeObject, fill = "wages")
```


# Contiguous 48 states

Filter to only the contiuous 48 statses + D.C.

```{r}
contiguous_states <- HelperShapeObject %>% 
  filter(REGION != 9) %>% 
  filter(STUSPS != "AK") %>% 
  filter(STUSPS != "HI")
  
```


## Make Choropleth

```{r}
tm_shape(contiguous_states) +
  tm_polygons("wages", id = "Name")
```


## Projection

Mark likes the USA_Contiguous_Albers_Equal_Area_Conic_USGS_version projection for the continental US. `EPSG:5070`

```{r}
contiguous_states %>% 
  st_transform(crs = "+init=epsg:5070") %>% 
  tm_shape() +
  tm_polygons("wages", id = "Name")
```