---
title: "some work with geom_sf and coordinates -- formerly TidyCensus"
---


Goal:  Make a choropleth with census data

- define choropleth
- define polygons
- define census geography
- Briefly Introduce ACS and decennial census 

    - census key 


```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidycensus)
library(leaflet)
library(sf)
library(stringr)
```


## Create a Simple Features dataframe using TidyCensus
```{r load-data, message=FALSE, warning=FALSE}
nc_pop <- 
  get_acs(geography = "county",
          variables = "B01003_001",
          state = "NC",
          geometry = TRUE)

as_tibble(nc_pop)
```




## Make Choropleth in Leaflet

Generate a color palette for the colors and legend

```{r colorpalette}
MapPalette <- colorQuantile(palette = "viridis", domain = nc_pop$estimate, n = 10)
```

## Leaflet Example

Make choropleth by filling county polygons (census geography) with correlated value (population variable *B01003_001*)

`st_transform()` - converts coordinate system of simple features

```{r make_choropleth}
nc_pop %>%
    st_transform(crs = "+init=epsg:4326") %>%
    leaflet(width = "100%") %>%
    addProviderTiles(provider = "CartoDB.Positron") %>%
    addPolygons(popup = ~ str_extract(NAME, "^([^,]*)"),
                stroke = FALSE,
                smoothFactor = 0,
                fillOpacity = 0.7,
                color = ~ MapPalette(estimate)) %>%
    addLegend("bottomright", 
              pal = MapPalette, 
              values = ~ estimate,
              title = "Population percentiles",
              opacity = 1)
```

## ggplot::geom_sf & ggmap

```{r message=FALSE, warning=FALSE}
library(ggmap)
library(ggplot2)
```

###  Get ggmap

```{r get_ggmap}
nc_via_ggmap <- get_map("burlington, NC", 
                        source = "google", 
                        maptype = "roadmap", 
                        zoom = 7)
ggmap(nc_via_ggmap)
```

### ggplot::geom_sf


```{r ggplot_sf}

nc_pop %>%
  ggplot(aes(fill = estimate, color = estimate)) +
  #geom_sf(show.legend = TRUE) +
  geom_sf() +
  coord_sf(crs = 3632) +
  # coord_sf(crs = st_crs(nc))
  theme(panel.background = element_rect(fill = "transparent"),
        panel.grid.major = element_line(color = "transparent"),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank()) #+
  #guides(fill = guide_legend(title = "foo"))

```

## Riffing 1

From: https://blog.dominodatalab.com/geographic-visualization-with-rs-ggmaps/


```{r}

nc_center = as.numeric(geocode("North Carolina"))

nc_ggmap <- ggmap(get_googlemap(center = nc_center, 
                                scale=2, 
                                zoom=6), 
                  extent="normal")

nc_ggmap
```


## Riff 2

### Getmap Raster Map from ggmap

```{r}
nc_basemap <- get_map(location="Burlington, NC", zoom=7, maptype = 'satellite')

ggmap(nc_basemap)
```

### Add a layer

```{r}

#ggmap(nc_basemap) +
#  ggplot(aes(fill = estimate, color = estimate)) +
#  geom_sf

# aes(x=long, lat, group = group, fill = cut_number(HOMIC_R, 5))

nc_wgs84 <-  nc_pop %>%
  ggplot(aes(fill = estimate, color = estimate)) +
  geom_sf() +
  coord_sf(crs = 4326)

nc_wgs84

#ggmap(nc_basemap) +
#  geom_polygon(data = nc_wgs84,
#               x = long,
#               y = lat,
#               group = group,
#               fill = estimate) +
  #ggplot(data = nc_wgs84,
  #       aes(fill = estimate, 
  #           color = estimate)
  #       ) +
 # geom_sf()


```

