---
title: "TidyCensus"
---

## Load Libraries
```{r}
library(tidyverse)
library(tidycensus)
library(leaflet)
library(sf)
library(stringr)
```

## Choose a census variable

I recommend using one of the following variables.  However, you can use the `tidycensus::load_variables()` and follow the "Searching for variables" [instructions](https://walkerke.github.io/tidycensus/articles/basic-usage.html#searching-for-variables). 

- `B19013_001E` - median household income
- `B08131_001E`	- AGGREGATE TRAVEL TIME TO WORK (IN MINUTES) OF WORKERS BY PLACE OF WORK--STATE AND COUNTY LEVEL: Worked in State of residence: Workers 16 years and over who did not work at home -- (Estimate)
- `B08103_001E` - MEDIAN AGE BY MEANS OF TRANSPORTATION TO WORK: Total: Taxicab, motorcycle, bicycle, or other means: Workers 16 years and over -- (Estimate)


## get_acs()

Load the variable and assign and object name for some USA county using the `get_acs` function.

```{r}
census_variable <- 
  get_acs(geography = "county",
          variables = "B08103_001E",
          state = "NC",
          geometry = TRUE)

```


```{r}
census_variable
```



## Make Choropleth in Leaflet


### Generate a Palette
Using the *estimate* variable in the census_variable sf object, generate a MapPalette.

```{r}
MapPalette <- colorBin(palette = "viridis",
                       domain = census_variable$estimate,
                       bins = 4,
                       pretty = TRUE,
                       reverse = TRUE)
```


### Make Leaflet choropleth

Make choropleth by filling county polygons (census geography) with correlated value (from the ACS)


```{r}
census_variable %>%
    st_transform(crs = "+init=epsg:4326") %>%
    leaflet(width = "100%") %>%
    addProviderTiles(provider = "CartoDB.Positron") %>%
    addPolygons(popup = ~ str_extract(NAME, "^([^,]*)"),
                stroke = FALSE,
                smoothFactor = 0,
                fillOpacity = 0.7,
                color = ~ MapPalette(estimate)) %>%
    addLegend("bottomright", 
              pal = MapPalette, 
              values = ~ estimate,
              title = "Median Age of<br>Travellers by<br>Taxi/Moto/Bike",
              opacity = 1)
```


## For 'B08131_001E'

AGGREGATE TRAVEL TIME TO WORK (IN MINUTES)

```{r}
census_variable <- 
  get_acs(geography = "county",
          variables = "B08131_001E",
          state = "NC",
          geometry = TRUE)
```

AGGREGATE TRAVEL TIME TO WORK (IN MINUTES)

Convert Minutes to hours, hours to months.

```{r}
census_variable <-  census_variable %>% 
  rename(estimate_original = estimate) %>% 
  mutate(estimate = estimate_original / 60 / 730.0008)

```


```{r}
MapPalette <- colorBin(palette = "viridis",
                       domain = census_variable$estimate,
                       bins = 7,
                       pretty = TRUE,
                       reverse = TRUE)
```

```{r}
census_variable %>%
    st_transform(crs = "+init=epsg:4326") %>%
    leaflet(width = "100%") %>%
    addProviderTiles(provider = "CartoDB.Positron") %>%
    addPolygons(popup = ~ str_extract(NAME, "^([^,]*)"),
                stroke = FALSE,
                smoothFactor = 0,
                fillOpacity = 0.7,
                color = ~ MapPalette(estimate)) %>%
    addLegend("bottomright", 
              pal = MapPalette, 
              values = ~ estimate,
              title = "Agregate Travel Time<br>in Months",
              opacity = 1)
```



## B19013_001E - median household income

```{r}
census_variable <- 
  get_acs(geography = "county",
          variables = "B19013_001E",
          state = "NC",
          geometry = TRUE)
```

```{r}
census_variable
```


```{r}
MapPalette <- colorNumeric(palette = "viridis",
                           domain = census_variable$estimate,
                           reverse = TRUE)
```


```{r}
census_variable %>%
    st_transform(crs = "+init=epsg:4326") %>%
    leaflet(width = "100%") %>%
    addProviderTiles(provider = "CartoDB.Positron") %>%
    addPolygons(popup = ~ str_extract(NAME, "^([^,]*)"),
                stroke = FALSE,
                smoothFactor = 0,
                fillOpacity = 0.7,
                color = ~ MapPalette(estimate)) %>%
    addLegend("bottomright", 
              pal = MapPalette, 
              values = ~ estimate,
              title = "Median Household Income",
              labFormat = labelFormat(prefix = "$"),
              opacity = 1)
```

<!-- -->