---
title: "Exercise 2: Choropleths with tidycensus"
---

## Load Libraries
```{r}
library(tidyverse)
library(tidycensus)
library(leaflet)
library(sf)
library(stringr)
```

## Set API key

Insert your own API Key.  You may request a free key from the Census API tool https://api.census.gov/data/key_signup.html

```{r}
# census_api_key("long-string-of-numbers-from-census-api-key-generator")
```


## Choose a census variable

I recommend using one of the following variables.  However, you can use the `tidycensus::load_variables()` and follow the "Searching for variables" [instructions](https://walkerke.github.io/tidycensus/articles/basic-usage.html#searching-for-variables). 

- `B19013_001E` - median household income
- `B08131_001E`	- AGGREGATE TRAVEL TIME TO WORK (IN MINUTES) OF WORKERS BY PLACE OF WORK--STATE AND COUNTY LEVEL: Worked in State of residence: Workers 16 years and over who did not work at home -- (Estimate)
- `B08103_001E` - MEDIAN AGE BY MEANS OF TRANSPORTATION TO WORK: Total: Taxicab, motorcycle, bicycle, or other means: Workers 16 years and over -- (Estimate)


## get_acs()

Load the variable and assign and object name for some USA county using the `get_acs` function.

```{r}
census_variable <- 
  get_acs(geography = "",
          variables = "",
          state = "",
          geometry = TRUE)

```




## Make Choropleth in Leaflet


### Generate a Palette
Using the *estimate* variable in the census_variable object, generate a MapPalette using the `leaflet::colorBin` function.

```{r}
MapPalette <- colorBin(palette = "viridis",
                       domain = ,
                       bins = )
```


### Make Leaflet choropleth

Make choropleth by filling county polygons (census geography) with correlated value (from the ACS)

1. Start with your "census_variable"
2. Set crs projection using `st_transform()` to EPSG "4326"
3. Pipe this to `leaflet()`
4. Assign the basemap with `addProviderTiles(provider = "CartoDB.Positron")`
5. using the `addPolygons()` function

    a. Assign a cleaned named to the popup window with `popup` argument and `= ~ str_extract(NAME, "^([^,]*)")`
    b. Assign the color palette with `color = ~ MapPalette(estimate))`
    c. use `stroke`, `smoothFactor`, and `fillOpacity` as you like
    
6. add a legend with `addLegend()`

```{r}
census_variable %>%
   
  
  
  
    
```


